---
title: "Assignment 4"
author: "Amisha Murarka"
date: "11/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install if necessary
#install.packages('gtools')
#load library
library(gtools)

```

## Scenerio 1

The World Values Survey is an ongoing worldwide survey that polls the world population about perceptions of life, work, family, politics, etc. The most recent phase of the survey that polled 77,882 people from 57 countries estimates that 36.2% of the world's population agrees with the statement "Men should have more right to a job than women." The survey also estimates that 13.8% of people have a university degree or higher and that 3.6% of people fit both criteria. 

#### Question 1: 
Are agreeing with the statement "Men should have more right to a job than women" and having a university degree or higher disjoint events?

##### Solution:
The given two events are not disjoint events as people who agree with the statement "Men should have more right to a job than women" can have a university degree also. The percentage estimates given are for the population and the respondants for both the events could overlap. We also know that probability of event A and event B is not 0 hence we cans say that they are not disjoint events.

| *Population - **77882***
| *Event A - Agreeing with the statement "Men should have more right to a job than women"*
| *A - **36.2%***
| *Event B - Having a university degree or higher*
| *B - **13.8%***
| *A & B - **3.6%***

#### Question 2: 
Draw a Venn diagram summarizing the variables and their associated probabilities.

##### Solution:
```{r}

#install.packages("VennDiagram")
library("VennDiagram")

pop <- 77882
a <- pop*0.362
pa <- a/pop # Probability of event A
b <- pop*0.138
pb <- b/pop # Probability of event B
both <- pop*0.036
pboth <- both/pop # Probability of A intersection B

# Plotting the Venn Diagram for both event probabilities
grid.newpage()  # Move to new plotting page
draw.pairwise.venn(area1 = pa,   # Create pairwise venn diagram
                   area2 = pb,
                   cross.area = pboth,
                   lty = "blank",
                   fill = c("pink", "green"),
                   category = c("A","B"))

```

#### Question 3: 
What is the probability that a randomly drawn person has a university degree or higher or agrees with the statement about men having more right to a job than women?

##### Solution:
0.464

```{r}

#Applying general addition rule to find probability of event A or B
p_a_or_b <- (pa+pb-pboth) 
p_a_or_b #Probability of a randomly drawn person having a university degree or higher or agreeing with the statement

```


#### Question 4: 
What percent of the world population do not have a university degree and disagree with the statement about men having more right to a job than women?

##### Solution:
53.6%

```{r}

#Probability of neither event A nor event B = 1 - Probability of event A or B
p_not_a_and_b <- 1-(p_a_or_b)
p_not_a_and_b*100 # Displaying the probability as percentage

```

#### Question 5: 
Does it appear that the event that someone agrees with the statement is independent of the event that they have a university degree or higher?

##### Solution:
Both the events are not independent as P(A and B) != P(A)*P(B)

```{r}

paste ("P(A and B) = ",pboth)
paste ("P(A)*P(B) = ",(pa*pb))

```


#### Question 6: 
What is the probability that at least 1 in 5 randomly selected people to agree with the statement about men having more right to a job than women?

##### Solution:
0.8942931

```{r}

P_A <- pa #Probability of a person agreeing to the statement
P_Not_A <- 1-P_A #Probability of a person not agreeing to the statement
#Pro of atleast one out of 5 agreeing to the statement = 1 - [Prob of not agreeing to the statement]^5
P_ATLEAST_ONE_AGREE <- 1-P_Not_A^5 
P_ATLEAST_ONE_AGREE

```

## Scenerio 2

As of 2009, Swaziland had the highest HIV prevalence in the world. 25.9% of this country's population is infected with HIV. The ELISA test is one of the first and most accurate tests for HIV. For those who carry HIV, the ELISA test is 99.7% accurate. For those who do not carry HIV, the test is 92.6% accurate. If an individual from Swaziland has tested positive, what is the probability that he carries HIV? Create a tree diagram to calculate the probability.

Prob(HIV) = 0.259
Prob(Not HIV) = 1-0.259 = 0.741
Prob(Positive|HIV) = 0.997 
Prob(Positive|Not HIV) = 0.926

#### Question 1: 
If an individual from Swaziland has tested positive, what is the probability that he carries HIV?

##### Solution:
0.8248434

```{r}

p_HIV<-0.259
p_Not_HIV<-1-p_HIV
p_Positive_Given_HIV <- 0.997
p_Negative_Given_Not_HIV <- 0.926
p_Positive_Given_Not_HIV <- 1-p_Negative_Given_Not_HIV
p_Negative_Given_HIV <- 1-p_Positive_Given_HIV

P_HIV_AND_Positive <- p_HIV * p_Positive_Given_HIV 
P_HIV_AND_Negative <- p_HIV*p_Negative_Given_HIV

P_Not_HIV_AND_Negative <- p_Not_HIV * p_Negative_Given_Not_HIV 
P_Not_HIV_AND_Positive <- p_Not_HIV*p_Positive_Given_Not_HIV

p_Positive <- P_HIV_AND_Positive + P_Not_HIV_AND_Positive 
P_HIV_GIVEN_Positive <- P_HIV_AND_Positive / p_Positive
P_HIV_GIVEN_Positive

```


#### Question 2: 
According to a 2013 Gallup poll, worldwide only 13% of employees are engaged at work (psychologically committed to their jobs and likely to be making positive contributions to their organizations). Among a random sample of 10 employees, what is the probability that 8 of them are engaged at work?

##### Solution:
2.77842e-06

```{r}

# p_Engaged <- 0.13
# p_Not_Engaged <- 1-p_Engaged
# 
# p_8_Engaged <- nrow(combinations(n=10,r=8)) * p_Engaged^8 * p_Not_Engaged^2
# p_8_Engaged 

# Probability of 8 employees out of 10 being engaged at work
dbinom(8, size = 10, p = 0.13)

#Plotting the binomial distribution
plot(0:10, dbinom(0:10, size = 10, p = 0.13),type='h',main='Binomial Distribution (n=245, p=0.25)',
     ylab='Probability',
     xlab ='Successes',)

```


#### Question 3: 
Recent study: “Facebook users get more than they give”

friend requests: 40% made, 63% received at least one
likes: liked 14 times, had their content “liked” 20 times, on average
messages: sent 9 messages, received 12, on average
 tags:12% tagged a friend in a photo, but 35% tagged other findings:
25% considered power users
average Facebook user has 245 friends
P(70 or more power user friends) = ?

##### Solution:
0.112763

```{r}

#Probability of having 70 or more power user friends
sum(dbinom(70:245, size = 245, p = 0.25))
#The probability value calculated using the above function is an approxiamted/normalised value. the exact value when calculated using the manual method or Z- score method comes out to be 0.098 

#Plotting the binomial distribution
plot(0:245, dbinom(0:245, size = 245, p = 0.25),type='h',main='Binomial Distribution (n=245, p=0.25)',
     ylab='Probability',
     xlab ='Successes',)

```


#### Question 4: 
According to a 2014 Gallup poll, 56% of uninsured Americans who plan to get health insurance say they will do so through a government health insurance exchange. What is the probability that in a random sample of 10 people exactly 6 plan to get health insurance through a government health insurance exchange?

##### Solution:
0.2427494

```{r}

#Probability of exactly 6 out of 10 people plan to get a health insurance through a government health insurance exchange
dbinom(6, size = 10, p = 0.56)

#Plotting the binomial distribution
plot(0:10, dbinom(0:10, size = 10, p = 0.56),type='h',main='Binomial Distribution (n=10, p=0.56)',
     ylab='Probability',
     xlab ='Successes',)

```

